---
title: "machine learning"
output: html_document
---

In this report, I try to predict the manner in which people did the exercise by analyzing the data from the Weight Lifting Exercise Dataset. After all the analyses, I use the prediction model to predict 20 different test cases.


```{r}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
set.seed(2048)
options(warn=-1)
```

Step1, Load the data as well as replace "#DIV/0!" with NA value

```{r}
training_data <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings=c("#DIV/0!") )
evaluation_data <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings=c("#DIV/0!") )
```

Step2, Set all columns 8 to the end to be numeric

```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}
```

Step3, Remove the columns which are almost blank and the columns of user name, timestamps and windows. Create a feature set which only contains the complete columns. 

```{r}
feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```

We now have the model data built from our feature set.

```{r}
idx <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```

Step4, Build 5 random forests with 150 trees each by paralell processing.

```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

Step5, Provide error reports for both training and test data to see how the model goes against the cross-validation set. As expected, the accuracy is around 99%.

```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```


Step6, Submission

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```